\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}

### 2023-10-03 manuscript/manuscriptRW.Rmd
changes following the review

The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter. Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality in mixed-species forests compared to even-aged, single-species forests. Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).

@@

The EuFoRIa data as used here are unique and well suited to better understand tree regeneration. However, three aspects of these data may represent considerable limitations. First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state. The data were collected in forest reserves where no management has taken place for long periods of time. This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight. In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher regeneration rates than the ones observed in the EuFoRia data. Thus, the overestimation of regeneration rates by the models may not be so problematic. Second, the data were collected for regeneration above a 7 cm threshold, thus limiting the assessment of tree regeneration to a holistic viewpoint. This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001]. In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess. Yet, regeneration data with lower thresholds are simply not available in a harmonized manner across large environmental gradients. Third, the empirical data were collected from rather small plots, while we sampled simulated regeneration levels from 1 ha areas, which may lead to an incorrect representation of space. Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022]. It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results. Ultimately, we need to acknowledge that the observed data represent just a snapshot of an extremely complex and stochastic reality. This may at least partly explain e.g. the lack of clear patterns in the simulation results along the gradients of climate water balance and degree-days.

@@

Therefore, we further recommend that more effort should be invested into collecting harmonized datasets on tree regeneration. We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future). Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g. the impacts of anthropogenic climate change.

>>>>

The fact that some models (and observed data) featured lower mortality than expected by Reineke's self-thinning rule may point to facilitation, or simply a higher mortality rate before the trees had reached 7 cm diameter. Yet, the case of models such as xComp, PICUS and TreeMig, whose mortality was well above the self-thinning line, indicates that they feature higher regeneration mortality compared to even-aged, single-species forests. Some models yielded a mortality rate of regeneration that is broadly compatible with the self-thinning rule (e.g., 4C and LandClim).

@@

The EuFoRIa data as used here are unique and well suited to better understand tree regeneration. However, three aspects of these data may represent considerable limitations. First, we made a comparison of tree regeneration in an equilibrium state, but we cannot assess how close the forests included in the EuFoRIa dataset were to such a state. The data were collected in forest reserves where no management has taken place for long periods of time. This makes our assumption of an equilibrium between forest properties and environmental drivers more reasonable than it might appear at first sight. In an analysis of primeval forests at demographic equilibrium, @brzeziecki2021 found higher regeneration rates than the ones observed in the EuFoRia data. Thus, the overestimation of regeneration rates by the models may not be so problematic. Second, the data were collected for regeneration above a 7 cm threshold, thus limiting the assessment of tree regeneration to a holistic viewpoint. This constituted a hard limit based on which we can understand only some aspects of tree regeneration, which in its entirety often comprises a rather long period since seed production [@price2001]. In reality, many environmental constraints are acting on young trees [@käber2021] that we were unable to assess. Yet, regeneration data with lower thresholds are simply not available in a harmonized manner across large environmental gradients. Third, the empirical data were collected from rather small plots, while we sampled simulated regeneration levels from 1 ha areas, which may lead to an incorrect representation of space. Even though the strategy we adopted is not ideal, it represents a common challenge when harmonizing diverse data sources originating from varied sampling strategies [@portier2022]. It would have been extremely challenging for such a variety of models to follow a protocol where the spatial sampling size was different at each of the 200 sites, and it would have introduced additional uncertainty in the results.

@@

Therefore, we further recommend that more effort should be invested into collecting harmonized datasets in a site-specific manner covering the different aspects leading to tree regeneration. We emphasize that datasets such as EuFoRIa are invaluable and should be expanded in both their spatial extent (e.g., towards boreal and Mediterranean conditions) as well as in time (e.g., continuing the monitoring into the future). Such data will allow for a better evaluation of forest models and help to reduce the uncertainty in their projections, which is crucial when they are used as tools for predicting e.g. the impacts of anthropogenic climate change.
